---
title: "hw2draft"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
```


### Problem 0

This solution focuses on a reproducible report containing code and text necessary for Problems 1-3, and is organized as an R Project. This was not prepared as a GitHub repo; examples for repository structure and git commits should be familiar from other elements of the course.

Throughout, we use appropriate text to describe our code and results, and use clear styling to ensure code is readable. 

```{r load_libraries}
library(tidyverse)
library(readxl)
```


### Problem 1

Below we import and clean data from `NYC_Transit_Subway_Entrance_And_Exit_Data.csv`. The process begins with data import, updates variable names, and selects the columns that will be used in later parts fo this problem. We update `entry` from `yes` / `no` to a logical variable. As part of data import, we specify that `Route` columns 8-11 should be character for consistency with 1-7.

```{r}
trans_ent = 
  read_csv(
    "data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

As it stands, these data are not "tidy": route number should be a variable, as should route. That is, to obtain a tidy dataset we would need to convert `route` variables from wide to long format. This will be useful when focusing on specific routes, but may not be necessary when considering questions that focus on station-level variables. 
7
The following code chunk selects station name and line, and then uses `distinct()` to obtain all unique combinations. As a result, the number of rows in this dataset is the number of unique stations.

```{r}
trans_ent %>% 
  select(station_name, line) %>% 
  distinct
```

* There are 465 distinct stations.

The next code chunk is similar, but filters according to ADA compliance as an initial step. This produces a dataframe in which the number of rows is the number of ADA compliant stations. 

```{r}
trans_ent %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

* There are 84 distinct stations, that are ADA compliant.

To compute the proportion of station entrances / exits without vending allow entrance, we first exclude station entrances that do not allow vending. Then, we focus on the `entry` variable -- this logical, so taking the mean will produce the desired proportion (recall that R will coerce logical to numeric in cases like this).

```{r}
trans_ent %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

* the proportion of station entrances / exits without vending allow entrance is 0.38.

Lastly, we write a code chunk to identify stations that serve the A train, and to assess how many of these are ADA compliant. As a first step, we tidy the data as alluded to previously; that is, we convert `route` from wide to long format. After this step, we can use tools from previous parts of the question (filtering to focus on the A train, and on ADA compliance; selecting and using `distinct` to obtain dataframes with the required stations in rows).

```{r}
trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

* There are 60 distinct stations, that serve the A train.
* There are 17 distinct stations, that serve the A train and are ADA compliant.


### Problem 2

```{r Mr Trash}
trash_W1 = 
  readxl::read_excel("data/Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N549") %>%
  janitor::clean_names() %>%
mutate (
        sports_balls = as.integer(sports_balls),
        wheel = "mrtrash")
```

Now, I will repeat the same for Professor data

```{r Prof Trash}
prof_W2 = 
  read_excel("data/Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", range = "A2:M96") %>%
  janitor::clean_names() %>%
mutate (year = as.character(year),
       wheel = "proftrash" )
```

Combining the both data sets:
```{r combining}
merged_trash = bind_rows (trash_W1, prof_W2) %>%
  janitor::clean_names()
```

Desription of the dataset post merging:

`r skimr:: skim(merged_trash)`

- The total number of rows are `r nrow(merged_trash)` and the total number of variables are `r ncol(merged_trash)`
- There are 3 character variables, 11 numeric variables and 1 date variable 
- The main variables in the dataset are weight of the trash (in tons), volume of the trash (in cubic yards) and the amount of various types of trash collected like plastic bottles, chips bags, cigarette butts, glass bottles, sports balls, polystyrene), the homes powered, and the distribution by names (Mr and Prof Trash).
- The total weight of trash collected combined is `r sum(merged_trash$weight_tons, na.rm = TRUE)` tons
- The total weight of trash collected by Professor Trash is `r sum(prof_W2$weight_tons, na.rm = TRUE)` tons
- The total number of sports balls collected by Mr Trash is `r sum(merged_trash[which(merged_trash$year == "2020" & merged_trash$wheel == "mrtrash"), "sports_balls"])`


### Problem 3

```{r polsmonth}
pols_ds = read_csv(
    "data/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(col=mon, into = c("year", "month", "day"), sep ='-', convert = TRUE) %>%
  mutate(month = month.abb[month],
         president = case_when (prez_gop == 1 ~ "gop", prez_dem == 1 ~ "dem")) %>%
  select(-prez_gop, -prez_dem, -day)
```

```{r snp}
snp_ds = read_csv(
    "data/fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  separate(col= date, into = c("month", "day", "year"), sep ='/', convert = TRUE) %>%
  mutate(month = month.abb[month],
         year = ifelse(year > 49, year + 1900, year + 2000)) %>%
select(-day) %>%
  select (year, month, everything())

```

```{r unemployment}
unemploy_ds = read_csv(
    "data/fivethirtyeight_datasets/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer (jan:dec,
names_to = "month",
values_to = "unemployment")
  
```

```{r}
pomo_snp = left_join(pols_ds, snp_ds)
```

```{r}
pomosnp_unemp = left_join(pomo_snp,unemploy_ds)
```

